General constraints on magic state protocols from the statistical mechanics of \\Wigner negativity

### 2021-06-25 magic_paper.tex
Edits done on my side. Refs to be added, and final proofreading before sending off.

\begin{abstract}
Magic states are key ingredients in schemes to realise universal fault-tolerant quantum computation.
Theories of magic states attempt to quantify this computational element via monotones and determine how these states may be efficiently transformed into useful forms. Here, we develop a statistical mechanical framework based on majorization in which Wigner negative magic states provide non-classical free energy that is processed under Clifford circuits for odd prime qudits. While traditional free energy is not defined in the presence of negativity, we show that majorization allows us to both quantify disorder in the Wigner representation and derive magic distillation bounds. These upper bounds are shown to be tighter than other bounds, such as from mana and thauma, and can incorporate hardware physics, such as temperature dependence and physical Hamiltonians. We also show that a subset of single-shot R\'{e}nyi entropies remain well-defined on quasi-distributions, are fully meaningful in terms of data processing and can acquire negative values that signal magic. We show that the mana of a magic state is the measure of divergence of these R\'{e}nyi entropies as one approaches the Shannon entropy for Wigner distributions, and discuss how distillation lower bounds could be obtained in this setting. This use of majorization for quasi-distributions could find application in other studies of non-classicality, and raises novel questions in the context of classical statistical mechanics.
\end{abstract}

@@

\begin{theorem} Majorization in any $\R_\sigma$ implies the monotonicity of sum-negativity. 
\end{theorem}
\begin{proof}
	The sum-negativity of a magic state $\rho$ can be written as $sn(\rho) =\frac{1}{2} (\sum_{\bmz} |W_\rho(\bmz) | - 1)$.
We make use of the $L_1$-norm form of relative majorization, (see~\cref{app:major}), which states that $\p \succ_{\r} \q$ if and only if
	\begin{equation}
\sum_k | p_k - r_k t | \geq \sum_k | q_k - r_k t |,
\end{equation}
for all $t\in \mathbb{R}$. Choosing $t=0$ we get the single condition that $\sum_k |p_k| \ge |\sum_k |q_k|$, independent of the choice of $\r$. Applying this to the Wigner quasi-distributions of two quantum states immediately gives the result.
\end{proof}

If we have a magic state $\rho$ that has negativity in its Wigner representation, then, as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. There is an simple relation between $L_\star$ and sum-negativity or mana, which is provided by the following result. 
\begin{theorem}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\R_\sigma$ and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every $\R_\sigma$.
\end{theorem}
The proof of this is given in~\cref{app:major}. Using the Lorenz curve perspective, it is also simple to construct new monotones. For example, within a given $\R_\sigma$, the area above the horizontal line $y = 1$ is another magic monotone.
\begin{lemma}
Given a magic state $\rho$ and a free state $\sigma$, let $\A_\sigma(\rho)$ be the area of the region $\{(x, y): 1 \leq y \leq L_{\rho | \sigma}(x)\}$. Then $\A_\sigma(\rho)$ is a magic monotone for $\R_\sigma$.
\end{lemma}
\begin{proof}
Consider the transformation $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$. We have that $L_{\rho_1|\sigma}(x)$ is never below the curve $L_{\rho_2|\sigma}(x)$, and therefore the region above $L=1$ for $\rho_2$ is a subset of the corresponding region for $\rho_1$. Thus, $\A_\sigma(\rho_1) \geq \A_\sigma(\rho_2)$, so $\A_\sigma$ is a magic monotone in $\R_\sigma$.
\end{proof}
Note though, in contrast to mana, the area monotone is \emph{specific to the $\R_\sigma$}, and its value will vary as we change $\sigma$. Therefore its monotonicity depends on the physics of the fixed point and provides a means to analyse magic distillation only under free operations that leave $\sigma$ invariant. One can also define an area monotone in the more general setting of relative majorization, where one drops the fixed point emphasis as in Section \ref{sec:stab}, but we do not explore this here.

@@

We can go further, and show that in fact a range of R\'{e}nyi entropies remain both well-defined and meaningful in the Wigner representation. The $\alpha$--R\'{e}nyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha.
\end{equation}
To extend to quasi-distributions we must ensure the following: (a) the entropy is well-defined on quasi-distributions, and (b) obeys monotonicity under stabilizer/free operations. In the Appendices we show the following.
\begin{theorem}\label{thm:HSchur} 
	If $\alpha = \frac{2a}{2b-1}$ for non-negative integers $a,b$ with $a \geq b$, then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'$ then $H_\alpha (\w) \leq H_\alpha(\w')$.
\end{theorem}
This monotonicity provides a meaningful extension of entropies to quasi-distributions, and the allowed values of $\alpha$ are dense in $\alpha >1$. Moreover we can prove the following result, which establishes the equivalence between Wigner negativity and the existence of negative R\'{e}nyi entropy.
\begin{theorem}\label{thm:Magic}
	A quantum state $\rho$ has Wigner negativity if and only if $H_\alpha(W_\rho) <0$ for some $\alpha =  \frac{2a}{2b-1}$, with integers $a \ge b$.
\end{theorem}
Therefore, the statistical mechanical description of stabilizer operations provides a setting in which negative entropy is fully meaningful, and quantifies the fact that magic states are more ordered than a perfectly sharp, deterministic classical state $\w = (1,0, \dots , 0)$ with zero entropy. We note that prior work has shown that negative conditional entropy~\cite{rio_thermodynamic_2011} arises in the context of quantum correlations, however this is unrelated to the present negative entropy.

The $\alpha \rightarrow 1$ limit of the R\'{e}nyi entropy is not well-defined, however the R\'{e}nyi entropy description shows that this diverges in a well-defined way. We can write the divergence of the entropy as 
\begin{equation}
	H_{1+\epsilon}(W_\rho) = -\epsilon^{-1} \log \sum_{\bmz} |W_\rho (\bmz)|^{1+\epsilon}
\end{equation} 
with $\epsilon \rightarrow 0^+$ through a sequence of rational values. We now recall that mana can be written as $\M(\rho) = \log \sum_{\bmz} |W_\rho(\bmz)|$, and so we obtain a second new description of mana:
\begin{equation}\label{eq:mana-divergence}
\M(\rho) = -\lim_{\epsilon \rightarrow 0^+} \epsilon H_{1-\epsilon}(W_\rho).
\end{equation}
Therefore, the mana of a state $\rho$ is the residue of a pole in $H_\alpha(W_\rho)$, and quantifies \emph{the divergence of the R\'{e}nyi-entropy} as we approach the limiting Shannon entropy $H(W_\rho)$.

We note that the R\'{e}nyi entropy can be described as a $q$--deformation of the Shannon entropy~\cite{baez2011renyi}, and $H_\alpha(\p)$ can be related to the $q$--derivative of free energy.  For this, the R\'{e}nyi parameter becomes a temperature term $\alpha = \frac{T_0}{T}$ where $T_0$ is a reference temperature and derivatives are considered via the limit $T \rightarrow T_0$. Therefore, if we try to push this to the present setting, the presence of Wigner negativity would correspond to a divergence in the first derivative of an effective free energy as $T \rightarrow T_0$. We can therefore speculate whether the presence of magic in Clifford circuits admits a sensible interpretation in terms of a zeroth-order phase transition [\ddd{cite}]. However we do not pursue this speculative line here, but just note that other works exist in which computational transitions are considered in relation to phase transitions [\ddd{cite}].

\subsection{R\'{e}nyi divergences and general distillation bounds}
Every entropy can be obtained from a relative entropy~\cite{Gour_2020} and so we now state the extensions for R\'{e}nyi divergences of Wigner distributions.
The $\alpha$--R\'{e}nyi divergence $D_\alpha(\p||\r)$ is defined as
\begin{equation}
	D_\alpha(\p \hspace{1pt}||\hspace{1pt} \r) := \frac{1}{\alpha-1} \log \sum_i p_i^\alpha r_i^{1-\alpha},
\end{equation}
where $\p$ can be a quasi-distribution, $\r$ a probability distribution with positive components and $\alpha$ can take the same values as for $H_\alpha$.
In~\cref{app:lcst_technical}, we also show that the divergence satisfies the following.

\begin{theorem}\label{thm:Da_props} 
	Let $\tau$ be in the interior of $\F$. 
	If $\alpha = \frac{2a}{2b-1}$ for non-negative integers $a,b$ with $a \geq b$, then the $\alpha$-R\'{e}nyi divergence $D_\alpha(W_\rho || W_\tau)$ is well-defined for all states $\rho$, and the following hold:
\begin{enumerate}
\item $D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) = 0$ if and only if $\rho =\tau$.
\item $D_\alpha(W_{\rho^{\otimes n}} \hspace{1pt}||\hspace{1pt} W_{\tau^{\otimes n}}) = n D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)$ for any $n \in \mathbb{N}$.
\item $D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \geq D_\alpha (W_{\E(\rho)} \hspace{1pt}||\hspace{1pt} W_{\E(\tau)})$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}

We can now apply this to obtain magic distillation bounds that describe Wigner negativity directly as a form of non-classical free energy.
\begin{theorem}\label{thm:entropic_bounds} 
	Consider a magic state distillation protocol that converts a magic state $\rho^{\otimes n} \longrightarrow \rho'^{\otimes m}$, which when applied to a reference stabilizer state $\tau^{\otimes n}$ results in $\tau^{\otimes n} \longrightarrow \tau'^{\otimes m}$.

	We have the following family of upper bounds on the magic distillation rate $R := m/n$,
	\begin{equation}
		R \leq \frac{F_\alpha(W_\rho) - F_\alpha(W_\tau)}{F_\alpha(W_{\rho'}) - F_\alpha(W_{\tau'})},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(W_\rho) := D_\alpha (W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) - \log \Z$ is the $\alpha$--free energy of the Wigner distribution. \ddd{[$\Z$ needs discussion]}
\end{theorem}
\ddd{[Perhaps remove free energy form...and allow $\frac{1}{m} D_\alpha(\rho'_m|| \tau'_m)$ general term with no assumption on correlations. ]}

For magic protocols that give unital maps, the above bound reduces to the simpler form
\begin{equation}
	R \leq \frac{2\log d - H_{\alpha}(W_\rho)}{2\log d - H_{\alpha}(W_{\rho'})}.
\end{equation}
In particular for the case of noisy Strange states on qutrits and $\epsilon'=0$, this bound becomes
\begin{equation}
	R \leq \frac{2(1-\alpha)\log 3 - \log \left [8(\frac{1}{6} - \frac{\epsilon}{18})^\alpha + (-\frac{1}{3} + \frac{4\epsilon}{9})^\alpha\right ] }{2(1-\alpha)\log 3 + \alpha \log 6 - \log (8 + 2^\alpha )}.
\end{equation}
The upper bound can be numerically minimized over $\alpha$, for any $\epsilon$.
%
%\begin{figure}[t!]
%    \centering
%    \subfigure[][]{%
%    \label{fig:Ra}%
%    \includegraphics[scale=0.4]{figs/Ra_vs_a_eps.pdf}}\\
%    \subfigure[][]{%
%    \label{fig:RaDiff}%
%    \includegraphics[scale=0.4]{figs/RaDiff_vs_a_eps.pdf}}
%    \caption{$R_\alpha$ and $R_\alpha - R_{\infty}$ for the Strange state distillation for unital protocols
%    \nick{Can compare thauma and mana for other magic states now more easily}}
%    \label{fig:rate_a_contour}
%\end{figure}

@@

\section{Single-shot entropies on quasi-distributions and magic distillation bounds}

@@

We next show that $H_\alpha (\w)$ for $\alpha = 2a/(2b-1)$ is a Schur-concave function on $\mathcal{Q}_N$.

\begingroup
\def\thetheorem{\ref{thm:HSchur}}
\begin{theorem}
	If $\alpha = \frac{2a}{2b-1}$ for non-negative integers $a,b$ with $a \geq b$, then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions $\mathcal{Q}_N$, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'\in \mathcal{Q}_N$ then $H_\alpha (\w) \leq H_\alpha(\w')$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
We have $\alpha = \frac{2a}{2b-1} \geq \frac{2b}{2b-1} > 1$ and $\alpha$ is a rational with even numerator and odd denominator, so $\sum_i w_i^\alpha > 0$ and $H_\alpha(\w)$ is well-defined for all quasi-distributions $\w$.

@@

Therefore, whenever $w_i \geq w_j$, we have that 
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which implies that
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i}(\w) \leq \frac{\partial H_\alpha}{\partial w_j}(\w),
\end{equation}
for any $\w$ in the interior of $\mathcal{Q}_N^\downarrow$.
Therefore, $H_\alpha$ is Schur-concave on $\mathcal{Q}_N$ and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w' \in \mathcal{Q}_N$ that obey $\w \succ \w'$.
\end{proof}
We note that if $\alpha =a/b$ with both $a$ and $b$ odd, then $H_\alpha$ is not well-defined for all quasi-probabilities, however if the actual set of quasi-distributions has a sufficiently bounded negativity, then $\log \sum_{\bmz} W_\rho(\bmz)^\alpha$ can still be obtained for $a$ odd, provided the sum is never negative.

If $\alpha =2a/(2b-1)$ with $\alpha < 1$, then $f(w) := \frac{-w^{\alpha-1}}{\alpha-1}$ is not monotonic. For example, if $\alpha = 2/3$, then $f(-2/6) = -4.33$, $f(-1/6) = -5.45$ and $f(+2/6) = +4.33$.. \ddd{[Needs edits]}

  Finally, we note that the set $F := \{2a/(2b-1): a,b \in \mathbb{N} \text{ and } a \geq b\}$ is dense in the reals $\mathbb{R}_{>1}$, as any rational $c/d$ with $c>d$ can be approximated by  $c2^n / (d2^n-1) \in F$ for sufficiently large $n$

We also have that $H_{\alpha}$ is additive on products of quasi-distributions. 
\begin{lemma}\label{H_add}
	For any $\w \in \mathcal{Q}_N$ and $\w' \in \mathcal{Q}_{N'}$, we have
	\begin{equation}
		H_{\alpha}(\w\otimes \w') = H_{\alpha}(\w) + H_{\alpha}(\w'),
	\end{equation}
	where $\alpha = 2a/(2b-1)$ with integers $a\ge b$.
\end{lemma}
\begin{proof}
	\begin{align*}
		H_{\alpha}(\w \otimes \w') &= \frac{1}{1-\alpha}\log \sum_{i,j'} \left[\w_i \w'_j\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\log  \left [ \sum_i\w_i^\alpha \sum_j\w_j^{\prime \alpha} \right ] \\
				&= \frac{1}{1-\alpha}\log  \left [ \sum_i\w_i^\alpha\right ] + \frac{1}{1-\alpha}\log \left [ \sum_j\w_j^{\prime \alpha} \right ] \\
		&= H_{\alpha}(\w) + H_{\alpha}(\w').
	\end{align*}
\end{proof}

@@

\subsection{R\'{e}nyi divergences on quasi-distributions}
Now we define the classical $\alpha$-R\'{e}nyi divergence in terms of the Wigner quasi-distribution of any magic state $\rho$ and the Wigner probability distribution of any equilibrium state $\tau$ as
\begin{equation}
	D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) := \frac{1}{\alpha - 1} \log \sum_{\bmz} W_{\rho}(\bmz)^\alpha W_{\tau}(\bmz)^{1-\alpha},
\end{equation}
where $\alpha$ can take values of the form $2a/(2b+1)$ \ddd{[2b+1 or 2b-1??]}, for non-negative integers $a,b$ to ensure once again that the argument of the logarithm is positive.

We can connect the $\alpha$-R\'{e}nyi entropy and divergence for quasi-distributions via the following result.

>>>>

\begin{abstract}
Magic states are key ingredients in schemes to realise universal fault-tolerant quantum computation.
Theories of magic states attempt to quantify this computational element via monotones and determine how these states may be efficiently transformed into useful forms. Here, we develop a statistical mechanical framework based on majorization in which Wigner negative magic states provide non-classical free energy that is processed under Clifford circuits for qudits of odd prime dimension. While traditional free energy is not defined in the presence of negativity, we show that majorization allows us to both quantify disorder in the Wigner representation and derive magic distillation bounds. These upper bounds are shown to be tighter than other bounds, such as from mana and thauma, and can incorporate hardware physics, such as temperature dependence and physical Hamiltonians. We also show that a subset of single-shot R\'{e}nyi entropies remain well-defined on quasi-distributions, are fully meaningful in terms of data processing and can acquire negative values that signal magic. We find that the mana of a magic state is the measure of divergence of these R\'{e}nyi entropies as one approaches the Shannon entropy for Wigner distributions, and discuss how distillation lower bounds could be obtained in this setting. This use of majorization for quasi-distributions could find application in other studies of non-classicality, and raises novel questions in the context of classical statistical mechanics.
\end{abstract}

@@

\begin{theorem} Majorization in any $\R_\sigma$ implies the monotonicity of sum-negativity/mana. 
\end{theorem}
\begin{proof}
	The sum-negativity of a magic state $\rho$ can be written as $sn(\rho) =\frac{1}{2} (\sum_{\bmz} |W_\rho(\bmz) | - 1)$.
We make use of the $L_1$-norm form of relative majorization, (see~\cref{app:major}), which states that $\p \succ_{\r} \q$ if and only if
	\begin{equation}
\sum_k | p_k - r_k t | \geq \sum_k | q_k - r_k t |,
\end{equation}
for all $t\in \mathbb{R}$. Choosing $t=0$ we get the single condition that $\sum_k |p_k| \ge |\sum_k |q_k|$, independent of the choice of $\r$. Applying this to the Wigner quasi-distributions of two quantum states immediately gives the result.
\end{proof}

If we have a magic state $\rho$ that has negativity in its Wigner representation, then, as discussed, its Lorenz curve over-shoots $1$ and reaches a non-trivial maximum $L_\star$ that depends on the particular state. There is an simple relation between $L_\star$ and sum-negativity/mana, which is provided by the following result. 
\begin{theorem}\label{lem:lcmax}
	Given a magic state $\rho$, the maximum $L_\star$ of its Lorenz curve $\lc{\rho}{\sigma}(x)$ is independent of the $\R_\sigma$ and equal to $1+\sn{\rho}$. Moreover, the majorization constraint is stronger than mana in every $\R_\sigma$.
\end{theorem}
The proof of this is given in~\cref{app:major}. Using the Lorenz curve perspective, it is also simple to construct new monotones. For example, within a given $\R_\sigma$, the area above the horizontal line $y = 1$ is another magic monotone.
\begin{lemma}
Given a magic state $\rho$ and a free state $\sigma$, let $\A_\sigma(\rho)$ be the area of the region $\{(x, y): 1 \leq y \leq L_{\rho | \sigma}(x)\}$. Then $\A_\sigma(\rho)$ is a magic monotone for $\R_\sigma$.
\end{lemma}
\begin{proof}
Consider the transformation $\rho_1 \rightarrow \rho_2$ within $\R_\sigma$. We have that $L_{\rho_1|\sigma}(x)$ is never below the curve $L_{\rho_2|\sigma}(x)$, and therefore the region above $L=1$ for $\rho_2$ is a subset of the corresponding region for $\rho_1$. Thus, $\A_\sigma(\rho_1) \geq \A_\sigma(\rho_2)$, so $\A_\sigma$ is a magic monotone in $\R_\sigma$.
\end{proof}
Note though, in contrast to mana, the area monotone is \emph{specific to the $\R_\sigma$}, and its value will vary as we change $\sigma$. Therefore its monotonicity depends on the physics of the fixed point and provides a means to analyse magic distillation only under free operations that leave $\sigma$ invariant. One can also define an area monotone in the more general setting of relative majorization, where one drops the fixed point emphasis as in Section \ref{sec:stab}, but we do not explore this here.

@@

We can go further, and show that in fact a range of R\'{e}nyi entropies remain both well-defined and meaningful in the Wigner representation. The $\alpha$--R\'{e}nyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha.
\end{equation}
To extend to quasi-distributions we must ensure the following: (a) the entropy is well-defined on quasi-distributions, and (b) obeys monotonicity under the majorization pre-order. In Appendix~\ref{app:single-shot} we prove the following.
\begin{theorem}\label{thm:HSchur} 
	If $\alpha = \frac{2a}{2b-1}$ for positive integers $a,b$ with $a \geq b$, then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'$ then $H_\alpha (\w) \leq H_\alpha(\w')$.
\end{theorem}
This monotonicity provides a meaningful extension of entropies to quasi-distributions. The unusual form of the $\alpha$ parameter is required so that the entropy function is well-defined and real-valued for all $\w$. The allowed values of $\alpha$ are dense in $\alpha >1$, however, while R\'{e}nyi entropies on probabilities can also be used for $0 < \alpha <1$ it turns out that for quasi-probability distributions this parameter regime is no longer monotonic in terms of the majorization pre-order (see Appendix~\ref{app:single-shot} for details).

We can prove the following result, which establishes the equivalence between Wigner negativity and the existence of negative R\'{e}nyi entropy.
\begin{theorem}\label{thm:Magic}
	A quantum state $\rho$ has Wigner negativity if and only if $H_\alpha(W_\rho) <0$ for some $\alpha =  \frac{2a}{2b-1}$, with positive integers $a \ge b$.
\end{theorem}
Therefore, the statistical mechanical description of stabilizer operations provides a setting in which negative entropy is fully meaningful, and quantifies the fact that magic states are more ordered than a perfectly sharp, deterministic classical state $\w = (1,0, \dots , 0)$ with zero entropy. We note that prior work has shown that negative conditional entropy~\cite{rio_thermodynamic_2011} arises in the context of quantum correlations, however this is unrelated to the present negative entropy.

The $\alpha \rightarrow 1$ limit of the R\'{e}nyi entropy is not well-defined, however it diverges in a well-defined way. We can write the divergence of the entropy as 
\begin{equation}
	H_{1+\epsilon}(W_\rho) = -\epsilon^{-1} \log \sum_{\bmz} |W_\rho (\bmz)|^{1+\epsilon}
\end{equation} 
with $\epsilon \rightarrow 0^+$ through a sequence of rational values. We now recall that mana can be written as $\M(\rho) = \log \sum_{\bmz} |W_\rho(\bmz)|$, and so we obtain a second new description of mana:
\begin{equation}\label{eq:mana-divergence}
\M(\rho) = -\lim_{\epsilon \rightarrow 0^+} \epsilon H_{1+\epsilon}(W_\rho).
\end{equation}
Therefore, the mana of a state $\rho$ is the residue of a pole in $H_\alpha(W_\rho)$, and quantifies \emph{the divergence of the R\'{e}nyi-entropy} as we approach the limiting Shannon entropy $H(W_\rho)$.

We also note that the R\'{e}nyi entropy can be described as a $q$--deformation of the Shannon entropy~\cite{baez2011renyi}, and that $H_\alpha(\p)$ can be related to the $q$--derivative of free energy.  For this, the R\'{e}nyi parameter becomes a temperature term $\alpha = \frac{T_0}{T}$ where $T_0$ is a reference temperature and derivatives are considered via the limit $T \rightarrow T_0$. Therefore, if we try to push this to the present setting, the presence of Wigner negativity would correspond to a divergence in the first derivative of an effective free energy as $T \rightarrow T_0$. It is interesting to speculate whether the presence of magic in Clifford circuits could admit a sensible interpretation in terms of zeroth-order phase transitions [\ddd{cite}].

\subsection{R\'{e}nyi divergences and general distillation bounds}
Every entropy can be obtained from a relative entropy~\cite{Gour_2020} and so we now state the extensions for R\'{e}nyi divergences of Wigner distributions.
The $\alpha$--R\'{e}nyi divergence $D_\alpha(\w||\r)$ is defined as
\begin{equation}
	D_\alpha(\w \hspace{1pt}||\hspace{1pt} \r) := \frac{1}{\alpha-1} \log \sum_i w_i^\alpha r_i^{1-\alpha},
\end{equation}
where $\w$ can be a quasi-distribution, $\r$ a probability distribution with positive components. In~\cref{app:single-shot}, we also show that the divergence satisfies the following.

\begin{theorem}\label{thm:Da_props} 
	Let $\tau$ be in the interior of $\F$. 
	If $\alpha = \frac{2a}{2b-1}$ for positive integers $a,b$ with $a \geq b$, then the $\alpha$-R\'{e}nyi divergence $D_\alpha(W_\rho || W_\tau)$ is well-defined for all quantum states $\rho$, and the following hold:
\begin{enumerate}
\item $D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) = 0$ if and only if $\rho =\tau$.
\item $D_\alpha(W_{\rho^{\otimes n}} \hspace{1pt}||\hspace{1pt} W_{\tau^{\otimes n}}) = n D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)$ for any $n \in \mathbb{N}$.
\item $D_\alpha(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau) \geq D_\alpha (W_{\E(\rho)} \hspace{1pt}||\hspace{1pt} W_{\E(\tau)})$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}

We can now apply this to obtain a family of magic distillation bounds based on R\'{e}nyi divergences.
\begin{theorem}
	Consider a general magic state distillation process on prime dimension qudits, that converts a magic state $\rho^{\otimes n} \longrightarrow \E(\rho^{\otimes n})=\rho'^{\otimes m}$and let $\tau$ be any full-rank stabilizer reference state on a qudit. Then the distillation rate $R := m/n$ is upper bounded as,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho \hspace{1pt}||\hspace{1pt} W_\tau)}{D^{\rm reg}_\alpha( \rho', \tau')},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and with the regularized divergence
	\begin{equation}
D^{\rm reg}_\alpha( \rho', \tau') = \frac{1}{m} D_\alpha (W_{\rho'^{\otimes m}} || W_{\tau'_m}),
\end{equation}
between the output magic state $\rho'^{\otimes m}$ and $\tau'_m = \E(\tau^{\otimes n})$.
\end{theorem}
The proof is given in Appendix~\ref{app:single-shot}.

While this bound is abstract in its present form, it does suggest future applications by viewing $D^{\rm reg}_\alpha( \rho', \tau')$ within the context of hypothesis testing, as a measure of distinguishability between $\rho'^{\otimes m}$ and the $m$ qudit state $\tau'_m$. 
However in order to properly link with hypothesis-testing one would have to first extend such results to quasi-distributions.

If correlations in $\tau'_m$ between subsystems can be neglected then we simply get
\begin{equation}
D^{\rm reg}_\alpha( \rho', \tau') = D_\alpha (W_{\rho'} || W_{\tau'}),
\end{equation}
which gives a generalized version of Theorem \ref{thm:free-energy}.

We can illustrate a simple application of the above bound for magic protocols that give unital maps. In this case it is easy to see that the above bound reduces to the simple form
\begin{equation}
	R \leq R_\alpha=\frac{2\log d - H_{\alpha}(W_\rho)}{2\log d - H_{\alpha}(W_{\rho'})}.
\end{equation}
In particular for the case of noisy Strange states on qutrits and $\epsilon'=0$, this bound becomes
\begin{equation}
	R \leq R_\alpha=\frac{2(1-\alpha)\log 3 - \log \left [8(\frac{1}{6} - \frac{\epsilon}{18})^\alpha + (-\frac{1}{3} + \frac{4\epsilon}{9})^\alpha\right ] }{2(1-\alpha)\log 3 + \alpha \log 6 - \log (8 + 2^\alpha )}.
\end{equation}
This upper bound can be numerically minimized over $\alpha$, for any $\epsilon$. The case $R_{10}$ is shown in Figure~\ref{fig:distill_bounds} and numerical evidence shows it is close to the optimal bound from majorization.
%
%\begin{figure}[t!]
%    \centering
%    \subfigure[][]{%
%    \label{fig:Ra}%
%    \includegraphics[scale=0.4]{figs/Ra_vs_a_eps.pdf}}\\
%    \subfigure[][]{%
%    \label{fig:RaDiff}%
%    \includegraphics[scale=0.4]{figs/RaDiff_vs_a_eps.pdf}}
%    \caption{$R_\alpha$ and $R_\alpha - R_{\infty}$ for the Strange state distillation for unital protocols
%    \nick{Can compare thauma and mana for other magic states now more easily}}
%    \label{fig:rate_a_contour}
%\end{figure}

@@

\section{Single-shot entropies on quasi-distributions and magic distillation bounds}\label{app:single-shot}

@@

We next show that $H_\alpha (\w)$ for $\alpha = 2a/(2b-1)$ is a Schur-concave function on $\mathcal{Q}_N$ provided that $a \ge b$.

\begingroup
\def\thetheorem{\ref{thm:HSchur}}
\begin{theorem}
	If $\alpha = \frac{2a}{2b-1}$ for positive integers $a,b$ with $a \geq b$, then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions $\mathcal{Q}_N$, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'\in \mathcal{Q}_N$ then $H_\alpha (\w) \leq H_\alpha(\w')$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
We have $\alpha = \frac{2a}{2b-1} \geq \frac{2b}{2b-1} > 1$ and $\alpha$ is a rational with even numerator and odd denominator, so $\sum_i w_i^\alpha > 0$ and $H_\alpha(\w)$ is well-defined for all quasi-distributions $\w$.

@@

Therefore, whenever $w_i \geq w_j$, we have that 
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which implies that
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i}(\w) \leq \frac{\partial H_\alpha}{\partial w_j}(\w),
\end{equation}
for any $\w$ in the interior of $\mathcal{Q}_N^\downarrow$.
Therefore, $H_\alpha$ is Schur-concave on $\mathcal{Q}_N$ and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w' \in \mathcal{Q}_N$ that obey $\w \succ \w'$.
\end{proof}
While we have integers $a,b$ such that $0 < \alpha=\frac{2a}{2b-1} < 1$ and $H_\alpha(\w)$ is well-defined, it turns out that monotonicity does not hold if we drop the condition $a \ge b$. If $\alpha =2a/(2b-1)$ with $\alpha < 1$, then $g(w) := \frac{-w^{\alpha-1}}{\alpha-1}$ is no longer monotonic for all $w \in \mathbb{R}$, and the problem occurs when comparing $w_i < 0$ and $w_j >0$. For example the function $g(x) = \frac{1}{x^3}$ is monotone decreasing on $x<0$ and $x>0$ separately, however it is not monotone on the full real-line.

We note that if $\alpha =r/s$ with both $r$ and $s$ odd, then $H_\alpha$ is not well-defined for all quasi-probabilities, however if the actual set of quasi-distributions has a sufficiently bounded levels negativity, then $\log \sum_{\bmz} W_\rho(\bmz)^\alpha$ can still be obtained for $r$ odd, provided the total sum is never negative. This occurs for Wigner distributions, where we have $|W_\rho(\z)| \ge 1/d$ for a $d$--dimensional quantum system.

We also note that the set $F := \{2a/(2b-1): a,b \in \mathbb{N} \text{ and } a \geq b\}$ is dense in the reals $\mathbb{R}_{>1}$, as any rational $c/d$ with $c>d$ can be approximated by  $c2^n / (d2^n-1) \in F$ for sufficiently large $n$.

We have that $H_{\alpha}$ is additive on products of quasi-distributions., which we state for completeness.
\begin{lemma}\label{H_add}
	For any $\w \in \mathcal{Q}_N$ and $\w' \in \mathcal{Q}_{N'}$, we have
	\begin{equation}
		H_{\alpha}(\w\otimes \w') = H_{\alpha}(\w) + H_{\alpha}(\w'),
	\end{equation}
	where $\alpha = 2a/(2b-1)$ with positive integers $a\ge b$.
\end{lemma}
\begin{proof}
	\begin{align*}
		H_{\alpha}(\w \otimes \w') &= \frac{1}{1-\alpha}\log \sum_{i,j'} \left[\w_i \w'_j\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\log  \left [ \sum_i\w_i^\alpha \sum_j\w_j^{\prime \alpha} \right ] \\
				&= \frac{1}{1-\alpha}\log  \left [ \sum_i\w_i^\alpha\right ] + \frac{1}{1-\alpha}\log \left [ \sum_j\w_j^{\prime \alpha} \right ] \\
		&= H_{\alpha}(\w) + H_{\alpha}(\w').
	\end{align*}
\end{proof}

@@

\subsection{R\'{e}nyi divergences on quasi-distributions}
We now define the $\alpha$-R\'{e}nyi divergence for a quasi-distribution $\w \in \mathcal{Q}_N$ relative to a full-rank probability distribution $\r\in \mathcal{Q}_N^+$ as
\begin{equation}
	D_\alpha(\w||\r) := \frac{1}{\alpha - 1} \log \sum_{i} w_i^\alpha r_i^{1-\alpha},
\end{equation}
where $\alpha=2a/(2b-1)$ for positive integers $a \ge b$. We now have the following result that relates the R\'{e}nyi divergence to the R\'{e}nyi entropy on a dense subset of probability distributions.
\begin{theorem}\label{thm:H2D}
